Reading episode files and preparing data sets

### 2025-11-06 data-raw/01-preprocessing-episode-files.Rmd
Adjusting origins/destinations to include travel options (such as walking to take the bus or car)

|-------------|------------------|-------------|------------------------------|

@@

-   `PUMFID`: the record identification[^2].
-   `WGHT_EPI`: Episode weight.
-   `ACTCODE`: Activity code of the episode.
-   `STARTIME`: Start time of the episode.
-   `ENDTIME`: End time of the episode.
-   `DURATION` Duration (in minutes) of the episode
-   `LOCATION`: Location/mode of transport of the episode.
-   `ORIGIN`: Place of origin of the episode.
-   `DESTINATION`: Place of destination of the episode.

[^2]: This variable serves as the link between the Main and the Episode
    files. Each case has a unique record identifier and this appears on
    every episode.

@@

The file name pattern is as follows: `transportation mode`[^3] + `_` +
`year of survey` [^4]. This means that, for instance, the file
`walking_2015` refers to transportation mode table `walking` in the year
2015.

[^3]:  `walking` or `cycling`

[^4]:  whether 1986, 1992, 1998, 2005, 2010, 2015, or 2022. Note: There is no
    data on cycling for the 1986 GSS survey.

@@

Now, we'll select all the rows with walking episodes in the GSS survey,
selecting the columns that we are interested[^5], and already creating
the origin and destination variables:

[^5]: Specifically for the 2015 GSS survey, `TUI_01` refers to the
    activity code. We'll change the name of this column to be match the
    name of the activity code variable in the others surveys.

@@

```{r}
walking_2015 <- walking_2015 |>
  filter(origin == 300 | 
           origin == 301 | 
           origin == 302 | 
           origin == 303 | 
           origin == 304 |
           origin == 305 | 
           origin == 306 | 
           origin == 307 |
           origin == 308 |
           origin == 309 | 
           origin == 310 |
           origin == 311) |>
  mutate(orig_label =
           case_when(origin == 300 ~ "Home", 
                     origin == 301 ~ "Work or school",
                     origin == 302 ~ "Business",
                     origin == 303 ~ "Other's home",
                     origin == 304 ~ "In the neighbourhood",
                     origin == 305 ~ "Outdoors",
                     origin == 306 ~ "Grocery store, other stores or mall",
                     origin == 307 ~ "Library, museum or theatre",
                     origin == 308 ~ "Sports centre, field or arena",
                     origin == 309 ~ "Restaurant, bar or club",
                     origin == 310 ~ "Place of worship",
                     origin == 311 ~ "Medical, dental or other health clinic"),
         orig_label = factor(orig_label))
```

@@

```{r}
walking_2015 <- walking_2015 |>  
    filter(destination == 300 | 
           destination == 301 | 
           destination == 302 | 
           destination == 303 |
           destination == 304 |
           destination == 305 |
           destination == 306 | 
           destination == 307 |
           destination == 308 |
           destination == 309 | 
           destination == 310 |
           destination == 311)  |>
  # Label the destinations
  mutate(dest_label =
           case_when(destination == 300 ~ "Home", 
                     destination == 301 ~ "Work or school",
                     destination == 302 ~ "Business",
                     destination == 303 ~ "Other's home",
                     destination == 304 ~ "In the neighbourhood",
                     destination == 305 ~ "Outdoors",
                     destination == 306 ~ "Grocery store, other stores or mall",
                     destination == 307 ~ "Library, museum or theatre",
                     destination == 308 ~ "Sports centre, field or arena",
                     destination == 309 ~ "Restaurant, bar or club",
                     destination == 310 ~ "Place of worship",
                     destination == 311 ~ "Medical, dental or other health clinic"),
         dest_label = factor(dest_label))
```

@@

Now, let's select only the cases of respondents who live in a CMA or CA:
```{r}
gss_m_2015 <- gss_m_2015[,c('PUMFID','LUC_RST','PRV')]

@@

The person was at home doing personal care (code `27`), then went for
walk that lasted 15 minutes, came back home and spent 180 minutes
looking for work (code `9`). So, this was a recreational/leisure
trip![^6]

[^6]: `TUI_01` is an activity code of the episode (What were you doing
    at minute). The list of main activity codes (001-095) can be found
    in an appendix of the User Guide. According to this list, `27` is
    "personal care", `7` is "Transport to or from activity" and `9` is
    "Looking for work".

Label the activity ("7" is "Transport to or from activity") and drop the LOCATION because it is already coded as MODE:    
```{r}
walking_2015 <- walking_2015 |>
  mutate(act_label = "Transport to or from activity") |>
  dplyr::select(PUMFID, WGHT_EPI, ACTCODE, act_label, STARTIME, ENDTIME, DURATION, origin, destination, orig_label, dest_label, YEAR, MODE, LUC_RST, PRV, Pop_centre, Province)
```

@@

```{r}
cycling_2015 <- cycling_2015 |>
  filter(origin == 300 | 
           origin == 301 | 
           origin == 302 | 
           origin == 303 | 
           origin == 304 |
           origin == 305 | 
           origin == 306 | 
           origin == 307 |
           origin == 308 |
           origin == 309 | 
           origin == 310 |
           origin == 311) |>
  mutate(orig_label =
           case_when(origin == 300 ~ "Home", 
                     origin == 301 ~ "Work or school",
                     origin == 302 ~ "Business",
                     origin == 303 ~ "Other's home",
                     origin == 304 ~ "In the neighbourhood",
                     origin == 305 ~ "Outdoors",
                     origin == 306 ~ "Grocery store, other stores or mall",
                     origin == 308 ~ "Sports centre, field or arena",
                     origin == 307 ~ "Library, museum or theatre",
                     origin == 309 ~ "Restaurant, bar or club",
                     origin == 310 ~ "Place of worship",
                     origin == 311 ~ "Medical, dental or other health clinic"),
         orig_label = factor(orig_label))
```

@@

```{r}
# change destination and origins to text column
cycling_2015 <- cycling_2015 |>  
  # Choose destinations of interest
  filter(destination == 300 | 
           destination == 301 | 
           destination == 302 | 
           destination == 303 |
           destination == 304 |
           destination == 305 |
           destination == 306 | 
           destination == 307 |
           destination == 308 |
           destination == 309 | 
           destination == 310 |
           destination == 311)  |>
  # Label the destinations
  mutate(dest_label =
           case_when(destination == 300 ~ "Home", 
                     destination == 301 ~ "Work or school",
                     destination == 302 ~ "Business",
                     destination == 303 ~ "Other's home",
                     destination == 304 ~ "In the neighbourhood",
                     destination == 305 ~ "Outdoors",
                     destination == 306 ~ "Grocery store, other stores or mall",
                     destination == 307 ~ "Library, museum or theatre",
                     destination == 308 ~ "Sports centre, field or arena",
                     destination == 309 ~ "Restaurant, bar or club",
                     destination == 310 ~ "Place of worship",
                     destination == 311 ~ "Medical, dental or other health clinic"),
         dest_label = factor(dest_label))
```

@@

The individual was at home engaged in personal care activities
(`TUI_01` = `2`) before cycling for approximately 10 minutes to a grocery
store. After spending 30 minutes at the store, they cycled back home.
Upon returning, the person prepared a meal and cleaned the house for approximately 85 minutes. Subsequently, the individual embarked on
another bicycle trip to the store, which took 10 minutes, and thereafter
returned home.[^7]

[^7]: Note: `TUI_01` is an activity code of the episode (What were you
    doing in minutes?).The list of main activity codes (001-095) can be
    found in an appendix of the User Guide.

Label the activity ("7" is "Transport to or from activity") and drop the LOCATION because it is already coded as MODE:    
```{r}
cycling_2015 <- cycling_2015 |>
  mutate(act_label = "Transport to or from activity") |>
  dplyr::select(PUMFID, WGHT_EPI, ACTCODE, act_label, STARTIME, ENDTIME, DURATION, origin, destination, orig_label, dest_label, YEAR, MODE, LUC_RST, PRV, Pop_centre, Province)
```

Save data in R format for later use:
```{r}
usethis::use_data(cycling_2015, 
                  overwrite = TRUE)
```

@@

```{r}
walking_2010 <- walking_2010 |>
  filter(origin == 1 | 
           origin == 2 | 
           origin == 3 | 
           origin == 4 | 
           origin == 5 |
           origin == 6 | 
           origin == 7 | 
           origin == 8 |
           origin == 9 |
           origin == 10 ) |>
  mutate(orig_label =
           case_when(origin == 1 ~ "Home", 
                     origin == 2 ~ "Work or school",
                     origin == 3 ~ "Other's home",
                     origin == 4 ~ "Restaurant, bar or club",
                     origin == 5 ~ "Place of worship",
                     origin == 6 ~ "Grocery store, other stores or mall",
                     origin == 7 ~ "Grocery store, other stores or mall",                     
                     origin == 8 ~ "Work or school",
                     origin == 9 ~ "Outdoors",
                     origin == 10 ~ "Library, museum or theatre"),
         orig_label = factor(orig_label))
```

@@

```{r}
walking_2010 <- walking_2010 |>
  filter(destination == 1 | 
           destination == 2 | 
           destination == 3 | 
           destination == 4 | 
           destination == 5 |
           destination == 6 | 
           destination == 7 | 
           destination == 8 |
           destination == 9 |
           destination == 10 ) |>
  mutate(dest_label =
           case_when(destination == 1 ~ "Home", 
                     destination == 2 ~ "Work or school",
                     destination == 3 ~ "Other's home",
                     destination == 9 ~ "Outdoors",
                     destination == 6 ~ "Grocery store, other stores or mall",
                     destination == 10 ~ "Library, museum or theatre",
                     destination == 4 ~ "Restaurant, bar or club",
                     destination == 5 ~ "Place of worship",
                     destination == 7 ~ "Grocery store, other stores or mall",                     
                     destination == 8 ~ "Work or school"),
         dest_label = factor(dest_label))
```

@@

Now, let's select only the cases of respondents who live in a CMA or CA:
```{r}
gss_m_2010 <- gss_m_2010[,c('RECID','LUC_RST','PRV')]

@@

The individual spent their time at home engaging in meal preparation
(`ACTCODE` =`430`) for a duration of 15 minutes, followed by a walk
lasting approximately 20 minutes to reach a grocery store. After
spending 15 minutes there, the person walked an additional 10 minutes to
another grocery store, where they stayed for 10 minutes before returning
home in 20 minutes. Once home, the individual engaged in socialization
for approximately 10 minutes (`ACTCODE`=`751`) before taking a 10-minute
walk to another location. At this new location, the person undertook
activities related to washing and dressing (`ACTCODE`=`400`) for around 10
minutes, then went for a 10-minute walk, returning home after that.
Subsequently, they dedicated 10 minutes to meal preparation
(`ACTCODE`=`101`), followed by a 10-minute session of food (or meal) cleanup
(`ACTCODE`=`110`). The individual then took a 15-minute walk, returned home,
and spent 5 minutes on household management tasks, which included
organizing and planning activities.[^8]

@@

```{r}
cycling_2010 <- cycling_2010 |>
  filter(origin == 1 | 
           origin == 2 | 
           origin == 3 | 
           origin == 4 | 
           origin == 5 |
           origin == 6 | 
           origin == 7 | 
           origin == 8 |
           origin == 9 |
           origin == 10 ) |>
  mutate(orig_label =
           case_when(origin == 1 ~ "Home", 
                     origin == 2 ~ "Work or school",
                     origin == 3 ~ "Other's home",
                     origin == 9 ~ "Outdoors",
                     origin == 6 ~ "Grocery store, other stores or mall",
                     origin == 10 ~ "Library, museum or theatre",
                     origin == 4 ~ "Restaurant, bar or club",
                     origin == 5 ~ "Place of worship",
                     origin == 7 ~ "Grocery store, other stores or mall",                     
                     origin == 8 ~ "Work or school"),
         orig_label = factor(orig_label))
```

@@

```{r}
cycling_2010 <- cycling_2010 |>
  filter(destination == 1 | 
           destination == 2 | 
           destination == 3 | 
           destination == 4 | 
           destination == 5 |
           destination == 6 | 
           destination == 7 | 
           destination == 8 |
           destination == 9 |
           destination == 10 ) |>
  mutate(dest_label =
           case_when(destination == 1 ~ "Home", 
                     destination == 2 ~ "Work or school",
                     destination == 3 ~ "Other's home",
                     destination == 9 ~ "Outdoors",
                     destination == 6 ~ "Grocery store, other stores or mall",
                     destination == 10 ~ "Library, museum or theatre",
                     destination == 4 ~ "Restaurant, bar or club",
                     destination == 5 ~ "Place of worship",
                     destination == 7 ~ "Grocery store, other stores or mall",                     
                     destination == 8 ~ "Work or school"),
         dest_label = factor(dest_label))
```

@@

Save data in R format for later use:
```{r}
usethis::use_data(cycling_2010, 
                  overwrite = TRUE)
```

@@

```{r}
walking_2005 <- walking_2005 |>
  filter(origin == 1 | 
           origin == 2 | 
           origin == 3 | 
           origin == 4 | 
           origin == 5 |
           origin == 6 | 
           origin == 7 | 
           origin == 8 |
           origin == 9 |
           origin == 10 ) |>
  mutate(orig_label =
           case_when(origin == 1 ~ "Home", 
                     origin == 2 ~ "Work or school",
                     origin == 3 ~ "Other's home",
                     origin == 9 ~ "Outdoors",
                     origin == 6 ~ "Grocery store, other stores or mall",
                     origin == 10 ~ "Library, museum or theatre",
                     origin == 4 ~ "Restaurant, bar or club",
                     origin == 5 ~ "Place of worship",
                     origin == 7 ~ "Grocery store, other stores or mall",               
                     origin == 8 ~ "Work or school"),
         orig_label = factor(orig_label))
```

@@

```{r}
walking_2005 <- walking_2005 |>
  filter(destination == 1 | 
           destination == 2 | 
           destination == 3 | 
           destination == 4 | 
           destination == 5 |
           destination == 6 | 
           destination == 7 | 
           destination == 8 |
           destination == 9 |
           destination == 10 ) |>
  mutate(dest_label =
           case_when(destination == 1 ~ "Home", 
                     destination == 2 ~ "Work or school",
                     destination == 3 ~ "Other's home",
                     destination == 9 ~ "Outdoors",
                     destination == 6 ~ "Grocery store, other stores or mall",
                     destination == 10 ~ "Library, museum or theatre",
                     destination == 4 ~ "Restaurant, bar or club",
                     destination == 5 ~ "Place of worship",
                     destination == 7 ~ "Grocery store, other stores or mall",                     
                     destination == 8 ~ "Work or school"),
         dest_label = factor(dest_label))
```

@@

Now, let's select only the cases of respondents who live in a CMA or CA:
```{r}
gss_m_2005 <- gss_m_2005[,c('RECID','LUC_RST','PRV')]

@@

The individual engaged in personal grooming and dressing activities
(`ACTCODE`= 400) before embarking on a brief walk lasting approximately 5
minutes, after which they returned home. Subsequently, the person
participated in religious activities at their residence, which included
services, prayer, and Bible readings (`ACTCODE` = 640). Following these
activities, the individual undertook another walk of around 5 minutes,
returning home once more. At home, they engaged in various activities,
including socializing with friends and relatives and sleeping. Later,
the person walked for approximately 5 minutes to reach a different
location and then spent an additional 5 minutes walking to return
home.[^9]

[^9]: Note: `ACTCODE` is an activity code of the episode (What were you
    doing in minute).The list of main activity codes can be found in an
    appendix of the User Guide.

@@

Save data in R format for later use:
```{r}
usethis::use_data(walking_2005,
                  overwrite = TRUE)
```

@@

```{r}
cycling_2005 <- cycling_2005 |>
  filter(origin == 1 | 
           origin == 2 | 
           origin == 3 | 
           origin == 4 | 
           origin == 5 |
           origin == 6 | 
           origin == 7 | 
           origin == 8 |
           origin == 9 |
           origin == 10 ) |>
  mutate(orig_label =
           case_when(origin == 1 ~ "Home", 
                     origin == 2 ~ "Work or school",
                     origin == 3 ~ "Other's home",
                     origin == 9 ~ "Outdoors",
                     origin == 6 ~ "Grocery store, other stores or mall",
                     origin == 10 ~ "Library, museum or theatre",
                     origin == 4 ~ "Restaurant, bar or club",
                     origin == 5 ~ "Place of worship",
                     origin == 7 ~ "Grocery store, other stores or mall",                     
                     origin == 8 ~ "Work or school"),
         orig_label = factor(orig_label))
```

@@

```{r}
cycling_2005 <- cycling_2005 |>
  filter(destination == 1 | 
           destination == 2 | 
           destination == 3 | 
           destination == 4 | 
           destination == 5 |
           destination == 6 | 
           destination == 7 | 
           destination == 8 |
           destination == 9 |
           destination == 10 ) |>
  mutate(dest_label =
           case_when(destination == 1 ~ "Home", 
                     destination == 2 ~ "Work or school",
                     destination == 3 ~ "Other's home",
                     destination == 9 ~ "Outdoors",
                     destination == 6 ~ "Grocery store, other stores or mall",
                     destination == 10 ~ "Library, museum or theatre",
                     destination == 4 ~ "Restaurant, bar or club",
                     destination == 5 ~ "Place of worship",
                     destination == 7 ~ "Grocery store, other stores or mall",                     
                     destination == 8 ~ "Work or school"),
         dest_label = factor(dest_label))
```

@@

Save data in R format for later use:
```{r}
usethis::use_data(cycling_2005, 
                  overwrite = TRUE)
```
As we have already processed this year's data, we will clear the Episode and Main files for this year from the computer's memory before uploading any more data

@@

```{r}
walking_1998 <- walking_1998 |>
  filter(origin == 1 | 
           origin == 2 | 
           origin == 3 ) |>
  mutate(orig_label =
           case_when(origin == 1 ~ "Home", 
                     origin == 2 ~ "Work or school",
                     origin == 3 ~ "Other's home"),
         orig_label = factor(orig_label))
```

@@

>>>>

|----------------|----------------|----------------|--------------------------|

@@

-   `PUMFID`: the record identification[^1].
-   `WGHT_EPI`: Episode weight.
-   `ACTCODE`: Activity code of the episode.
-   `STARTIME`: Start time of the episode.
-   `ENDTIME`: End time of the episode.
-   `DURATION` Duration (in minutes) of the episode
-   `LOCATION`: Location/mode of transport of the episode.
-   `ORIGIN`: Place of origin of the episode.
-   `DESTINATION`: Place of destination of the episode.

[^1]: This variable serves as the link between the Main and the Episode
    files. Each case has a unique record identifier and this appears on
    every episode.

@@

The file name pattern is as follows: `transportation mode`[^2] + `_` +
`year of survey` [^3]. This means that, for instance, the file
`walking_2015` refers to transportation mode table `walking` in the year
2015.

[^2]: `walking` or `cycling`

[^3]: whether 1986, 1992, 1998, 2005, 2010, 2015, or 2022. Note: There
    is no data on cycling for the 1986 GSS survey.

@@

Now, we'll select all the rows with walking episodes in the GSS survey,
selecting the columns that we are interested[^4], and already creating
the origin and destination variables:

[^4]: Specifically for the 2022 GSS survey, `TUI_01` refers to the
    activity code. We'll change the name of this column to be match the
    name of the activity code variable in the others surveys.

@@

```{r}
walking_2015 <- walking_2015 |>
  filter(origin <= 321) |>
  mutate(orig_label =
           case_when(origin == 300 ~ "Home", 
                     origin == 301 ~ "Work or school",
                     origin == 302 ~ "Business",
                     origin == 303 ~ "Other's home",
                     origin == 304 ~ "In the neighbourhood",
                     origin == 305 ~ "Outdoors",
                     origin == 306 ~ "Grocery store, other stores or mall",
                     origin == 307 ~ "Library, museum or theatre",
                     origin == 308 ~ "Sports centre, field or arena",
                     origin == 309 ~ "Restaurant, bar or club",
                     origin == 310 ~ "Place of worship",
                     origin == 311 ~ "Medical, dental or other health clinic",
                     origin == 312 ~ "Elsewhere",
                     origin %in% c(313,314,319) ~ "Travel - Motorized",
                     origin %in% c(317,320,321) ~ "Travel - Other",
                     origin == 316 ~ "Travel - Transit",
                     origin == 315 ~ "Travel - Walk",
                     origin == 318 ~ "Travel - Bycicle"),
         orig_label = factor(orig_label))
```

@@

```{r}
walking_2015 <- walking_2015 |>  
  filter(destination <= 321) |>
  mutate(dest_label =
           case_when(destination == 300 ~ "Home", 
                     destination == 301 ~ "Work or school",
                     destination == 302 ~ "Business",
                     destination == 303 ~ "Other's home",
                     destination == 304 ~ "In the neighbourhood",
                     destination == 305 ~ "Outdoors",
                     destination == 306 ~ "Grocery store, other stores or mall",
                     destination == 307 ~ "Library, museum or theatre",
                     destination == 308 ~ "Sports centre, field or arena",
                     destination == 309 ~ "Restaurant, bar or club",
                     destination == 310 ~ "Place of worship",
                     destination == 311 ~ "Medical, dental or other health clinic",
                     destination == 312 ~ "Elsewhere",
                     destination %in% c(313,314,319) ~ "Travel - Motorized",
                     destination %in% c(317,320,321) ~ "Travel - Other",
                     destination == 316 ~ "Travel - Transit",
                     destination == 315 ~ "Travel - Walk",
                     destination == 318 ~ "Travel - Bycicle"),
         dest_label = factor(dest_label))
```

@@

Now, let's select only the cases of respondents who live in a CMA or CA:

```{r}
gss_m_2015 <- gss_m_2015[,c('PUMFID','LUC_RST','PRV')]

@@

The person was at home doing personal care (code `27`), then went for
walk that lasted 15 minutes, came back home and spent 180 minutes
looking for work (code `9`). So, this was a recreational/leisure
trip![^5]

[^5]: `TUI_01` is an activity code of the episode (What were you doing
    at minute). The list of main activity codes (001-095) can be found
    in an appendix of the User Guide. According to this list, `27` is
    "personal care", `7` is "Transport to or from activity" and `9` is
    "Looking for work".

Label the activity ("7" is "Transport to or from activity") and drop the
LOCATION because it is already coded as MODE:

```{r}
walking_2015 <- walking_2015 |>
  mutate(act_label = "Transport to or from activity") |>
  dplyr::select(PUMFID, WGHT_EPI, ACTCODE, act_label, STARTIME, ENDTIME, DURATION, origin, destination, orig_label, dest_label, YEAR, MODE, LUC_RST, PRV, Pop_centre, Province)
```

@@

```{r}
cycling_2015 <- cycling_2015 |>
  filter(origin <= 321) |>
  mutate(orig_label =
           case_when(origin == 300 ~ "Home", 
                     origin == 301 ~ "Work or school",
                     origin == 302 ~ "Business",
                     origin == 303 ~ "Other's home",
                     origin == 304 ~ "In the neighbourhood",
                     origin == 305 ~ "Outdoors",
                     origin == 306 ~ "Grocery store, other stores or mall",
                     origin == 307 ~ "Library, museum or theatre",
                     origin == 308 ~ "Sports centre, field or arena",
                     origin == 309 ~ "Restaurant, bar or club",
                     origin == 310 ~ "Place of worship",
                     origin == 311 ~ "Medical, dental or other health clinic",
                     origin == 312 ~ "Elsewhere",
                     origin %in% c(313,314,319) ~ "Travel - Motorized",
                     origin %in% c(317,320,321) ~ "Travel - Other",
                     origin == 316 ~ "Travel - Transit",
                     origin == 315 ~ "Travel - Walk",
                     origin == 318 ~ "Travel - Bycicle"),
         orig_label = factor(orig_label))
```

@@

```{r}
# change destination and origins to text column
cycling_2015 <- cycling_2015 |>  
  filter(destination <= 321) |>
  mutate(dest_label =
           case_when(destination == 300 ~ "Home", 
                     destination == 301 ~ "Work or school",
                     destination == 302 ~ "Business",
                     destination == 303 ~ "Other's home",
                     destination == 304 ~ "In the neighbourhood",
                     destination == 305 ~ "Outdoors",
                     destination == 306 ~ "Grocery store, other stores or mall",
                     destination == 307 ~ "Library, museum or theatre",
                     destination == 308 ~ "Sports centre, field or arena",
                     destination == 309 ~ "Restaurant, bar or club",
                     destination == 310 ~ "Place of worship",
                     destination == 311 ~ "Medical, dental or other health clinic",
                     destination == 312 ~ "Elsewhere",
                     destination %in% c(313,314,319) ~ "Travel - Motorized",
                     destination %in% c(317,320,321) ~ "Travel - Other",
                     destination == 316 ~ "Travel - Transit",
                     destination == 315 ~ "Travel - Walk",
                     destination == 318 ~ "Travel - Bycicle"),
         dest_label = factor(dest_label))
```

@@

The individual was at home engaged in personal care activities (`TUI_01`
= `2`) before cycling for approximately 10 minutes to a grocery store.
After spending 30 minutes at the store, they cycled back home. Upon
returning, the person prepared a meal and cleaned the house for
approximately 85 minutes. Subsequently, the individual embarked on
another bicycle trip to the store, which took 10 minutes, and thereafter
returned home.[^6]

[^6]: Note: `TUI_01` is an activity code of the episode (What were you
    doing in minutes?).The list of main activity codes (001-095) can be
    found in an appendix of the User Guide.

Label the activity ("7" is "Transport to or from activity") and drop the
LOCATION because it is already coded as MODE:

```{r}
cycling_2015 <- cycling_2015 |>
  mutate(act_label = "Transport to or from activity") |>
  dplyr::select(PUMFID, WGHT_EPI, ACTCODE, act_label, STARTIME, ENDTIME, DURATION, origin, destination, orig_label, dest_label, YEAR, MODE, LUC_RST, PRV, Pop_centre, Province)
```

```{r}
usethis::use_data(cycling_2015, 
                  overwrite = TRUE)
```

@@

```{r}
walking_2010 <- walking_2010 |>
  filter(origin <= 21 ) |>
  mutate(orig_label =
           case_when(origin == 1 ~ "Home", 
                     origin %in% c(2,8) ~ "Work or school",
                     origin == 3 ~ "Other's home",
                     origin == 9 ~ "Outdoors",
                     origin %in% c(6,7) ~ "Grocery store, other stores or mall",
                     origin == 10 ~ "Library, museum or theatre",
                     origin == 4 ~ "Restaurant, bar or club",
                     origin == 5 ~ "Place of worship",
                     origin == 11 ~ "Elsewhere",
                     origin %in% c(12,13,19) ~ "Travel - Motorized",
                     origin %in% c(18,20,21) ~ "Travel - Other",
                     origin %in% c(15,16) ~ "Travel - Transit",
                     origin == 14 ~ "Travel - Walk",
                     origin == 17 ~ "Travel - Bycicle"),
         orig_label = factor(orig_label))
```

@@

```{r}
walking_2010 <- walking_2010 |>
  filter(destination <= 21 ) |>
  mutate(dest_label =
           case_when(destination == 1 ~ "Home", 
                     destination %in% c(2,8) ~ "Work or school",
                     destination == 3 ~ "Other's home",
                     destination == 9 ~ "Outdoors",
                     destination %in% c(6,7) ~ "Grocery store, other stores or mall",
                     destination == 10 ~ "Library, museum or theatre",
                     destination == 4 ~ "Restaurant, bar or club",
                     destination == 5 ~ "Place of worship",
                     destination == 11 ~ "Elsewhere",
                     destination %in% c(12,13,19) ~ "Travel - Motorized",
                     destination %in% c(18,20,21) ~ "Travel - Other",
                     destination %in% c(15,16) ~ "Travel - Transit",
                     destination == 14 ~ "Travel - Walk",
                     destination == 17 ~ "Travel - Bycicle"),
         dest_label = factor(dest_label))
```

@@

Now, let's select only the cases of respondents who live in a CMA or CA:

```{r}
gss_m_2010 <- gss_m_2010[,c('RECID','LUC_RST','PRV')]

@@

The individual spent their time at home engaging in meal preparation
(`ACTCODE` =`430`) for a duration of 15 minutes, followed by a walk
lasting approximately 20 minutes to reach a grocery store. After
spending 15 minutes there, the person walked an additional 10 minutes to
another grocery store, where they stayed for 10 minutes before returning
home in 20 minutes. Once home, the individual engaged in socialization
for approximately 10 minutes (`ACTCODE`=`751`) before taking a 10-minute
walk to another location. At this new location, the person undertook
activities related to washing and dressing (`ACTCODE`=`400`) for around
10 minutes, then went for a 10-minute walk, returning home after that.
Subsequently, they dedicated 10 minutes to meal preparation
(`ACTCODE`=`101`), followed by a 10-minute session of food (or meal)
cleanup (`ACTCODE`=`110`). The individual then took a 15-minute walk,
returned home, and spent 5 minutes on household management tasks, which
included organizing and planning activities.[^7]

[^7]: Note: `ACTCODE` is an activity code of the episode (What were you
    doing in minute).The list of main activity codes can be found in an
    appendix of the User Guide.

@@

```{r}
cycling_2010 <- cycling_2010 |>
  filter(origin <= 21 ) |>
  mutate(orig_label =
           case_when(origin == 1 ~ "Home", 
                     origin %in% c(2,8) ~ "Work or school",
                     origin == 3 ~ "Other's home",
                     origin == 9 ~ "Outdoors",
                     origin %in% c(6,7) ~ "Grocery store, other stores or mall",
                     origin == 10 ~ "Library, museum or theatre",
                     origin == 4 ~ "Restaurant, bar or club",
                     origin == 5 ~ "Place of worship",
                     origin == 11 ~ "Elsewhere",
                     origin %in% c(12,13,19) ~ "Travel - Motorized",
                     origin %in% c(18,20,21) ~ "Travel - Other",
                     origin %in% c(15,16) ~ "Travel - Transit",
                     origin == 14 ~ "Travel - Walk",
                     origin == 17 ~ "Travel - Bycicle"),
         orig_label = factor(orig_label))
```

@@

```{r}
cycling_2010 <- cycling_2010 |>
  filter(destination <= 21 ) |>
  mutate(dest_label =
           case_when(destination == 1 ~ "Home", 
                     destination %in% c(2,8) ~ "Work or school",
                     destination == 3 ~ "Other's home",
                     destination == 9 ~ "Outdoors",
                     destination %in% c(6,7) ~ "Grocery store, other stores or mall",
                     destination == 10 ~ "Library, museum or theatre",
                     destination == 4 ~ "Restaurant, bar or club",
                     destination == 5 ~ "Place of worship",
                     destination == 11 ~ "Elsewhere",
                     destination %in% c(12,13,19) ~ "Travel - Motorized",
                     destination %in% c(18,20,21) ~ "Travel - Other",
                     destination %in% c(15,16) ~ "Travel - Transit",
                     destination == 14 ~ "Travel - Walk",
                     destination == 17 ~ "Travel - Bycicle"),
         dest_label = factor(dest_label))
```

@@

```{r}
usethis::use_data(cycling_2010, 
                  overwrite = TRUE)
```

@@

```{r}
walking_2005 <- walking_2005 |>
  filter(origin <= 21 ) |>
  mutate(orig_label =
           case_when(origin == 1 ~ "Home", 
                     origin %in% c(2,8) ~ "Work or school",
                     origin == 3 ~ "Other's home",
                     origin == 9 ~ "Outdoors",
                     origin %in% c(6,7) ~ "Grocery store, other stores or mall",
                     origin == 10 ~ "Library, museum or theatre",
                     origin == 4 ~ "Restaurant, bar or club",
                     origin == 5 ~ "Place of worship",
                     origin == 11 ~ "Elsewhere",
                     origin %in% c(12,13,19) ~ "Travel - Motorized",
                     origin %in% c(18,20,21) ~ "Travel - Other",
                     origin %in% c(15,16) ~ "Travel - Transit",
                     origin == 14 ~ "Travel - Walk",
                     origin == 17 ~ "Travel - Bycicle"),
         orig_label = factor(orig_label))
```

@@

```{r}
walking_2005 <- walking_2005 |>
  filter(destination <= 21 ) |>
  mutate(dest_label =
           case_when(destination == 1 ~ "Home", 
                     destination %in% c(2,8) ~ "Work or school",
                     destination == 3 ~ "Other's home",
                     destination == 9 ~ "Outdoors",
                     destination %in% c(6,7) ~ "Grocery store, other stores or mall",
                     destination == 10 ~ "Library, museum or theatre",
                     destination == 4 ~ "Restaurant, bar or club",
                     destination == 5 ~ "Place of worship",
                     destination == 11 ~ "Elsewhere",
                     destination %in% c(12,13,19) ~ "Travel - Motorized",
                     destination %in% c(18,20,21) ~ "Travel - Other",
                     destination %in% c(15,16) ~ "Travel - Transit",
                     destination == 14 ~ "Travel - Walk",
                     destination == 17 ~ "Travel - Bycicle"),
         dest_label = factor(dest_label))
```

@@

Now, let's select only the cases of respondents who live in a CMA or CA:

```{r}
gss_m_2005 <- gss_m_2005[,c('RECID','LUC_RST','PRV')]

@@

The individual engaged in personal grooming and dressing activities
(`ACTCODE`= 400) before embarking on a brief walk lasting approximately
5 minutes, after which they returned home. Subsequently, the person
participated in religious activities at their residence, which included
services, prayer, and Bible readings (`ACTCODE` = 640). Following these
activities, the individual undertook another walk of around 5 minutes,
returning home once more. At home, they engaged in various activities,
including socializing with friends and relatives and sleeping. Later,
the person walked for approximately 5 minutes to reach a different
location and then spent an additional 5 minutes walking to return
home.[^8]

@@

```{r}
usethis::use_data(walking_2005,
                  overwrite = TRUE)
```

@@

```{r}
cycling_2005 <- cycling_2005 |>
  filter(origin <= 21 ) |>
  mutate(orig_label =
           case_when(origin == 1 ~ "Home", 
                     origin %in% c(2,8) ~ "Work or school",
                     origin == 3 ~ "Other's home",
                     origin == 9 ~ "Outdoors",
                     origin %in% c(6,7) ~ "Grocery store, other stores or mall",
                     origin == 10 ~ "Library, museum or theatre",
                     origin == 4 ~ "Restaurant, bar or club",
                     origin == 5 ~ "Place of worship",
                     origin == 11 ~ "Elsewhere",
                     origin %in% c(12,13,19) ~ "Travel - Motorized",
                     origin %in% c(18,20,21) ~ "Travel - Other",
                     origin %in% c(15,16) ~ "Travel - Transit",
                     origin == 14 ~ "Travel - Walk",
                     origin == 17 ~ "Travel - Bycicle"),
         orig_label = factor(orig_label))
```

@@

```{r}
cycling_2005 <- cycling_2005 |>
  filter(destination <= 21 ) |>
  mutate(dest_label =
           case_when(destination == 1 ~ "Home", 
                     destination %in% c(2,8) ~ "Work or school",
                     destination == 3 ~ "Other's home",
                     destination == 9 ~ "Outdoors",
                     destination %in% c(6,7) ~ "Grocery store, other stores or mall",
                     destination == 10 ~ "Library, museum or theatre",
                     destination == 4 ~ "Restaurant, bar or club",
                     destination == 5 ~ "Place of worship",
                     destination == 11 ~ "Elsewhere",
                     destination %in% c(12,13,19) ~ "Travel - Motorized",
                     destination %in% c(18,20,21) ~ "Travel - Other",
                     destination %in% c(15,16) ~ "Travel - Transit",
                     destination == 14 ~ "Travel - Walk",
                     destination == 17 ~ "Travel - Bycicle"),
         dest_label = factor(dest_label))
```

@@

```{r}
usethis::use_data(cycling_2005, 
                  overwrite = TRUE)
```

@@

```{r}
walking_1998 <- walking_1998 |>
  filter(origin <= 10 ) |>
  mutate(orig_label =
           case_when(origin == 1 ~ "Home", 
                     origin == 2 ~ "Work or school",
                     origin == 3 ~ "Other's home",
                     origin == 4 ~ "Elsewhere",
                     origin %in% c(5,6) ~ "Travel - Motorized",
                     origin == 10 ~ "Travel - Other",
                     origin == 8  ~ "Travel - Transit",
                     origin == 7 ~ "Travel - Walk",
                     origin == 9 ~ "Travel - Bycicle"),
         orig_label = factor(orig_label))
```

@@

                               ACTCODE == 331 ~ "Financial Services (e.g. banking)", # p. 188
